%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fontsize=11pt]{scrartcl} % 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm, amssymb} % Math packages
\usepackage{mathrsfs}

\usepackage[margin=1in]{geometry}

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}

\newtheorem{lemma}{Lemma}
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{Topology} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Problem Set 5 \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Daniel Halmrast} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

\section*{Problem 1}
Show that for $(X,\mathscr{T})$ a compact Hausdorff space, any topology
$\mathscr{T}\subsetneq\mathscr{T}'$ makes $(X,\mathscr{T}')$ no longer compact
Hausdorff. Furthermore, for $\mathscr{T}'\subsetneq\mathscr{T}$, the space
$(X,\mathscr{T}')$ is not compact Hausdorff.
\\
\begin{proof}
    For this proof, we will show that for $\mathscr{T},\mathscr{T}'$ topologies
    on $X$ with $\mathscr{T}\subset\mathscr{T}'$ and the property that $X$ is
    compact Hausdorff under both these topologies, then they must be equal.

    To do so, we consider the identity function
    $id:(X,\mathscr{T}')\to(X,\mathscr{T})$, which is continuous since
    $\mathscr{T}\subset\mathscr{T}'$. Now, for any $C$ closed in
    $(X,\mathscr{T}')$, since $(X,\mathscr{T}')$ is compact and $id$ is
    continuous, $id(C)=C$ is compact in $(X,\mathscr{T})$. However, since
    $(X,\mathscr{T})$ is Hausdorff, $C$ must also be closed. Thus, $id$ is a
    closed map, and thus a homeomorphism, and the two topologies must be equal.
\end{proof}

\section*{Problem 2}
Prove that for compact subsets $A,B$ of $X$, their union $A\cup B$ is compact as
well.
\\
\begin{proof}
    Let $\mathscr{O}$ be an open cover of $A\cup B$. In particular,
    $\mathscr{O}$ covers $A$, and has a finite subset $\mathscr{O}_A$ that
    covers $A$. Similarly, there is a finite subset $\mathscr{O}_B$ that
    covers $B$. Their union $\mathscr{O}_A\cup\mathscr{O}_B$ covers $A\cup B$,
    and is finite, as it is the union of finite sets. It is also a subcover,
    since both $\mathscr{O}_A$ and $\mathscr{O}_B$ are subsets of
    $\mathscr{O}$. Thus. $\mathscr{O}$ has a finite subcover, and $A\cup B$ is
    compact as desired.
\end{proof}

\section*{Problem 3}
Suppose $A$ is a subset of a metric space. Show that $A$ is compact implies that
$A$ is closed and bounded. Give an example where the converse does not hold.
\\
\begin{proof}
    Suppose $A$ is a compact subset of a metric space $(X,d)$. Since all metric
    spaces are Hausdorff, it follows immediately that $A$ is closed (since it is
    a compact subset of a Hausdorff space).

    Now, fix $x\in A$, and consider the open cover $\mathscr{O} = \{V_n(x)\ |\
    n\in\N\}$ of balls centered at $x$ with radius $n$. This has a finite
    subcover, by compactness of $A$. So, let $N$ be the largest radius in the
    finite subcover. For each $V_m(x)$ in the finite subcover, then, we have
    that $V_m(x)\subset V_N(x)$. Thus, since the subcover covers $A$, it follows
    that $V_N(x)$ covers $A$, and thus $A$ is bounded.

    For an example where the converse fails, consider the metric space
    $(\R,d_{LA})$ of the real numbers with the discrete (Los Angeles) metric.
    The set $\R$ itself is closed and bounded (since everything is at most
    distance $1$ away from any particular point), but is clearly not compact,
    since it is an uncountable discrete set of points.
\end{proof}

\section*{Problem 4}
Show that for $X$ Hausdorff, $A,B$ disjoint compact subsets of $X$, then there
exist disjoint open sets $U$ and $V$ containing $A$ and $B$ respectively.
\\
\begin{proof}

\end{proof}

\end{document}
