%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fontsize=11pt]{scrartcl} % 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage{mathrsfs}
\usepackage{tikz-cd}

\usepackage[margin=1in]{geometry}

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\C}{\mathbb{C}}

\newcommand{\Hom}{\text{Hom}}

\newtheorem{lemma}{Lemma}
%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{Geometry} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Homework \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Daniel Halmrast} % Your name

\date{\normalsize\today} % Today's date or a custom date

\begin{document}

\maketitle % Print the title

% Problems
\section*{Problem 1}
Prove that $V^{**}\cong V$ for a finite dimensional vector space $V$.
\\
\\
\begin{proof}
    The isomorphism is given by
    \[
        \begin{aligned}
        \Phi:V\to V^{**}\\
        \Phi(v) = ev_v = (\phi\mapsto \phi(v))
    \end{aligned}
    \]
    First, we observe that this map is linear. Indeed, for $v_1,v_2\in V$ and
    $\alpha,\beta$ a scalar, we have
    \[
\begin{aligned}
    \Phi(\alpha v_1 + \beta v_2)(\phi) &= \phi(\alpha v_1 + \beta v_2)\\
    &=\alpha \phi(v_1) + \beta\phi(v_2)\\
    &=\alpha\Phi(v_1)(\phi) + \beta\Phi(v_2)(\phi)
\end{aligned}
    \]
    as desired.

    We need to show this is surjective and injective. Injectivity of $\Phi$ is
    easily shown by examining the kernel of $\Phi$. Suppose $v$ is such that
    $ev_v(\phi)= \phi(v)=0$ for all $\phi\in V^*$. Then, since $V^*$ separates points of
    $V$, it follows that $v=0$. Thus, the kernel is trivial, as desired.

    Now, we need to show this map is surjective. To do so, we appeal to $v$
    being finite-dimensional, and let $\{e_i\}$ be a basis for $V$ with dual
    basis $\{\omega^i\}$. Then, let $x\in V^{**}$. Define a corresponding vector
    $\tilde{x} = \sum_ix(\omega^i)e_i$. Then
    \[
\begin{aligned}
    \Phi(\tilde{x})(\phi) &= \phi(\tilde{x})\\
    &= \sum_i x(\omega^i)\phi(e_i)\\
    &= \sum_i x(\omega^i\phi(e_i))\\
    &= x(\phi)
\end{aligned}
    \]
    as desired. Thus, $\Phi$ is a linear isomorphism.
\end{proof}

\section*{Problem 2}
Prove that for $V$ a vector space with basis $v_i$ and dual basis $v^i$, the set
\[
    \{v^i\otimes v^j\ |\ 1\leq i,j\leq n\}
\]
forms a basis for $V^*\otimes V^*$.
\\
\\
\begin{proof}
    For this, we will show that the vector space
\[
    W = \text{span}(\{v^i\otimes v^j\ |\ 1\leq i,j\leq n\})
\]
satisfies the universal property of
    tensor products. That is, we wish to show that for each bilinear map
    $h:V^*\times V^*\to U$ for some vector space $U$, there is a unique linear
    map $\tilde{h}:W\to U$ such that the diagram
    \[
        \begin{tikzcd}
            V^*\times V^* \arrow[d, "\otimes"] \arrow[r, "h"] & U \\
            W \arrow[ru, "\tilde{h}", dotted] & 
        \end{tikzcd}
    \]
    commutes. We will guess that $\otimes:V^*\times V^*\to W$ is given as
    \[
        \otimes(a_iv^i,b_jv^j) = a_ib_jv^i\otimes v^j
    \]

    So, let $h:V^*\times V^*\to U$ be a bilinear map. Define $\tilde{h}:W\to U$
    as 
    \[
        \tilde{h}(\sum_{i,j}a_{ij}v^i\otimes v^j)
        =a_{ij}h(v^i,v^j)
    \]
    Then it is clear that the diagram
    \[
        \begin{tikzcd}
            V^*\times V^* \arrow[d, "\otimes"] \arrow[r, "h"] & U \\
            W \arrow[ru, "\tilde{h}", dotted] & 
        \end{tikzcd}
    \]
    commutes, since
    \[
\begin{aligned}
    h(a_iv^i,b_jv^j) &= a_ib_jh(v^i,v^j)\\
    \tilde{h}\circ\otimes(a_iv^i,b_jv^j) &= \tilde{h}(a_ib_jv^i\otimes v^j)\\
    &= a_ib_jh(v^i,v^j)
\end{aligned}
    \]
    It should be clear from construction that $\tilde{h}$ is unique.

    Thus, since every bilinear map from $V^*\times v^*$ factors through $W$, $W$
    satisfies the universal property of tensor products, and is isomorphic to
    $V^*\otimes V^*$. Thus, the set $\{v^i\otimes v^j\}$ forms a basis for
    $V^*\otimes V^*$ as desired.
\end{proof}

\section*{Problem 3}
Show that $dvol = \wedge_i\omega^i = \sqrt{|g|}dx^n$.
\\
\\
\begin{proof}
    Recall that $dx^n = \wedge_{i=1}^ndx^i$, and our manifold is
    $n$-dimensional.

    Recall that for an $n$-fold wedge product $\wedge_{i=1}^n v^i$, we have (for
    $v^i = a_j^i\omega^j$ in an orthonormal frame)
    \[
        \wedge_{i=1}^n v^i = |\det(a_j^1,\dots,a_j^n)|\wedge_{i=1}^n\omega^i
        =\sqrt{\det(A^tA)}\wedge_{i=1}^n\omega^i
    \]
    for $A$ the matrix with columns $a^i$. We can apply this to $v^i = dx^i$ to
    get the desired result.
\end{proof}

\section*{Problem 4}
Show that the definition of the integral of a top degree form on a single chart
is independent of choice of coordinates.
\\
\\
\begin{proof}
    Recall the definition of the integral of a top degree differential form on a
    compact set $K$ in a single coordinate frame is
    \[
        \int_K\omega = \int_{\phi(K)}f\circ\phi^{-1}dx^n
    \]
    for $\omega = fdx^1\wedge\dots\wedge dx^n$.

    We also have the change-of-coordinates formula for a diffeomorphism
    $F:\Omega_1\to\Omega_2$ as
    \[
        \int_{\Omega_2}fdy^n = \int_{\Omega_1}f\circ F|J_F|dx^n
    \]
    where $J_F$ is the jacobian of $F$.

    So, let $\phi:M\to\R^n$ be the original coordinate system, and let
    $\psi:M\to\R^n$ be another coordinate system covering $K$. Then, we have the
    diffeomorphism $F = \psi\circ\phi^{-1}$ and we can apply this to get
    \[
        \int_{\psi(K)}g\circ psi^{-1}dy^n = \int_{\phi(K)} g\circ\psi^{-1}\circ F|J_F|dx^n
    \]
    which is just
    \[
        \begin{aligned}
            \int_{\phi(K)} g\circ\psi^{-1}\circ F|J_F|dx^n &= \int_{\phi(K)}
            g\circ \psi^{-1}\circ 
            \psi\circ\phi^{-1}|J_F|dx^n\\
            &= \int_{\phi(K)} g\circ \phi^{-1} |J_F|dx^n\\
            &= \int_K g|J_F|dx^n\\
            &= \int_K\omega
        \end{aligned}
    \]
    where we used the fact that $\omega = g dy^1\wedge\dots\wedge dy^n =
    g|J_F|dx^1\wedge\dots\wedge dx^n$ since $|J_F|=\det(J_F)$ on a two
    positively oriented charts.

    Thus, the two integrals agree.
\end{proof}

\section*{Problem 5}
Prove that a manifold is orientable if and only if it admits a nowhere vanishing
top degree form.
\\
\\
\begin{proof}
    Suppose $M$ is an orientable manifold. That is, there exists an atlas
    $\{U_{\alpha},\phi_{\alpha}\}$ of $M$ for which the Jacobian of each
    transition map has positive determinant. Let $\{\psi_{\alpha}\}$ be a
    partition of unity subordinate to the atlas $U_\alpha$. Then, define
    \[
        \omega = \sum_{\alpha}\psi_{\alpha}dx_{\alpha}^i\wedge\dots\wedge
        dx_{\alpha}^n
    \]
    where $x_{\alpha}^i$ are the coordinate functions on $U_\alpha$. We claim
    that $\omega$ is a nowhere-vanishing form. Clearly, if $p$ is a point in $M$
    contained in only one chart, then $\omega_p = dx_p^1\wedge\dots\wedge
    dx_p^n$ and does not vanish. If $p$ is such that it is contained in more
    than one chart, then $\omega$ at $p$ is the sum of positive terms (since
        each coordinate system is positive, we have $dx^n = \det(J)dy^n$ and
    $\det(J)$ is always positive) and does not vanish.
    \\
    \\
    Suppose instead that $M$ admits a nowhere-vanishing top degree form
    $\omega$. Then, let $\{U_{\alpha},\phi_{\alpha}\}$ be an atlas of $M$. For
    $x_{\alpha}^i$ the coordinate functions for $\phi_{\alpha}$, define a new
    coordinate system to be such that if $\omega$ is expressed as
    \[
    \omega = fdx_{\alpha}^1\wedge\dots\wedge dx_{\alpha}^n
    \]
    then $f$ is positive. This is done by setting $x_{\alpha}^1$ to its negative
    if $f$ is negative on that chart. Note that since $\omega$ never vanishes,
    we know that $f$ will be either entirely positive or entirely negative on a
    chart. Thus, such a choice can be made consistently.

    Then, the modified atlas is a positive coordinate chart for $M$, which is
    easily verified, since
    \[
    \omega = fdx^n = f\det(J)dy^n
    \]
    and $\omega$ always has positive coefficient, so $\det(J)$ is positive.
\end{proof}

\section*{Problem 6}
Show that the topology of $M$ coincides with the metric topology 
\[
    d_g(x,y) = \inf_{\gamma\in C^{\infty}(I,M)} \{L(\gamma)\ |\
    \gamma(0)=x,\gamma(1)=y\}
\]
where $L(\gamma)$ is the total length of $\gamma$ defined by
\[
    L(\gamma) = \int_Ig(\gamma',\gamma')dt
\]

\begin{proof}
    First, we observe the following: for $g$ a Riemannian metric, and $\gamma$ a
    curve contained entirely in a single coordinate chart $\phi$, there exist
    constants $c,C$ such that
    \[
        cL_{\R^n}(\gamma) \leq L_g(\gamma) \leq CL_{\R^n}(\gamma)
    \]
    where $L_{\R^n}(\gamma)$ is the length of $\phi\circ\gamma$ using the
    euclidean metric on $\R^n$.  This follows from the fact that the
    metric induced by $g$ along $\phi^{-1}$ defines a norm on $\R^n$, and all
    norms are equivalent. That is,
    \[
        k\|v\|_{\R^n} \leq \|v\|_{\phi^{-1*}g}\leq K\|v\|_{\R^n}
    \]
    for constants $k,K$. Thus, the lengths (defined in terms of integrals of the
    metric) follow the same inequality. 

    It should also be clear that the metrics induced by $\R^n$ and $g$ are
    equivalent as well. To see this, note that for any $x,y\in \R^n$, we have
    \[
        \begin{aligned}
        cd_{\R^n}(x,y) &= \inf_{\gamma(0)=x,\gamma(1)=y}cL_{\R^n}(\gamma)\\
        &\leq \inf_{\gamma(0)=x,\gamma(1)=y}L_g(\gamma)\\
        &=d_g(x,y)
    \end{aligned}
    \]
    (the first equality is proved in the next problem)
    and similarly for $d_g(x,y)\leq Cd_{\R^n}(x,y)$. This shows that the two
    topologies induced by the two metrics are equal.

    Equivalence of the two topologies follows immediately. We can show that for
    $U$ open in the manifold topology, and $x\in U$, there is a neighborhood $V$
    of $x$ in the metric topology contained in $U$. Simply take a coordinate
    ball $V_{\varepsilon}(x)$ of small enough radius to be contained in a single
    coordinate chart $\phi$. That is, the domain of $\phi$ contains
    $V_{\varepsilon}(x)$. Then, we know from the above observation that
    $\phi(V_{\varepsilon}(x))$ is open in $\R^n$ in the standard topology, and
    thus is open with respect to the pullback of $d_g$ along $\phi^{-1}$. Thus,
    $V_{\varepsilon}(x)$ is also open in the metric topology induced by $d_g$ on
    $M$. The same argument with the two topologies switched completes the
    argument that both topologies are equal.
\end{proof}

\section*{Problem 7}
Show that $\|a-b\|_{\R^n}$ is $d_{\R^n}(a,b) = \inf_{\gamma}L_{\R^n}(\gamma)$.
\\
\\
\begin{proof}
    This result follows from standard variational calculus on the functional
    $L_{\R^n}(\gamma)$. 

    Let's minimize the functional $L(\gamma)$ by varying the path $\gamma$. We
    do this by setting the variation to zero. That is, $L(\gamma)$ is maximized
    for $\gamma$ that makes $\delta L=0$. We calculate
    \[
\begin{aligned}
    \delta L &= \int \delta(\sqrt(g(\gamma',\gamma')))dt\\
    &=\int \frac{1}{2\sqrt{g(\gamma',\gamma')}}\delta g(\gamma',\gamma')dt
\end{aligned}
    \]
    Since arc length is independent of parameterization, we can take the unit
    speed parameterization of $\gamma$, so that $g(\gamma',\gamma')=0$. Then, we
    have
    \[
        \begin{aligned}
            \delta L &= \int \delta g(\gamma',\gamma')dt\\
            &= \int \delta (g_{ab}\partial_t\gamma^a\partial_t\gamma^b)dt\\
            &=\int g_{ab}\delta(\partial_t\gamma^a)\partial_t\gamma^b +
            g_{ab}\delta(\partial_t\gamma^b)\partial_t\gamma^a dt
            &= 2\int g_{ab}\partial_t(\delta\gamma^a)\partial_t\gamma^b dt
        \end{aligned}
    \]
    integrating by parts (and tossing boundary terms since the endpoints of
    $\gamma$ do not vary) and noting $g_{ab} = \delta_{ab}$ in $\R^n$ yields
    \[
        \delta L = -2\int \partial_t^2\gamma^a\delta\gamma^bdt
    \]
    which holds only if $\partial_t^2\gamma^a = 0$ for all $a$. Thus, the
    minimal length path from a point $p$ to a point $q$ is the straight line
    from $p$ to $q$.

    So, for $\gamma$ the straight line from $p$ to $q$, $L(\gamma) = \|\gamma\|
    = \|p-q\|$ as desired.
\end{proof}

\section*{Problem 8}
Define the Levi-Civita connection as the unique connection such that
\[
    \nabla_X Y-\nabla_YX = [X,Y]
\]
and
\[
    Xg(Y,Z) = g(\nabla_XY,Z) + g(Y,\nabla_XZ)
\]
Show that this is indeed a connection on $M$.
\\
\\
\begin{proof}
    For this proof, we will denote the Levi-Civita connection as $D$.

    We need to show that $D_XY$ is function linear in $X$, scalar linear in
    $Y$, and satisfies the Leibniz rule
    \[
        D_X(fY) = (Xf)Y + fD_XY
    \]

    Recall from class that by utilizing the two properties above, we see that
    \[
        2g(D_XY,Z) = Xg(Y,Z) + Yg(Z,X) - Zg(X,Y) + g([X,Y],Z) -g([Y,Z],X) -
        g([X,Z],Y)
    \]
    which uniquely determines the connection. Now, we just need to show that
    this definition satisfies the definition of a connection. To that end, let
    $f\in C^{\infty}(M)$. We calculate
    \[
\begin{aligned}
    2g(D_{fX}Y,Z) &= fXg(Y,Z) + Yg(Z,fX) - Zg(fX,Y)\\
        &+ g([fX,Y],Z) -g([Y,Z],fX) - g([fX,Z],Y)\\
        &=fXg(Y,Z) + Yg(Z,fX) - Zg(fX,Y)\\
        &- g([Y,fX],Z) - g([Y,Z],fX) + g([Z,fX],Y)\\
        &=fXg(Y,Z) + (Yf)g(Z,X) + fYg(Z,X) - (Zf)g(X,Y) - fZg(X,Y)\\
        &- g((Yf)X,Z) - g(f[Y,X],Z) - g([Y,Z],fX) + g((Zf)X,Y) + g(f[Z,X],Y)\\
        &=fXg(Y,Z)  + (Yf)g(Z,X) + fYg(Z,X) - (Zf)g(X,Y) - fZg(X,Y)\\
        &- g((Yf)X,Z) + g(f[X,Y],Z) - g([Y,Z],fX) + g((Zf)X,Y) - g(f[X,Z],Y)\\
        &=(fX)g(Y,Z) + (Yf)g(X,Z) - (Yf)g(X,Z) - (Zf)g(X,Y) + (Zf)g(X,Y)\\
        &+ fYg(X,Z) -fZg(X,Y) + g(f[X,Y],Z) -g([Y,Z],fX) -g(f[X,Z],Y)\\
        &= f\left\{ Xg(Y,Z) + Yg(X,Z) -Zg(X,Y) + g([X,Y],Z) - g([Y,Z],X) -
        g([X,Z],Y) \right\}\\
        &=fg(D_XY,Z) = g(fD_XY,Z)
\end{aligned}
    \]
    and thus $D$ is $C^{\infty}(M)$-linear in $X$.
\end{proof}<++>

\end{document}
